---
title: "Data Tidying"
author: "Dave Bosworth"
date: "11/6/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
# Load packages
library(tidyverse)
```

# Read and Clean Data

```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"), stringsAsFactors = FALSE)

head(catch_original)
```

Remove `all` and `notes` column because we don't need them.

```{r}
catch_data <- catch_original %>% 
  select(-c(All, notesRegCode))

head(catch_data)
```

```{r}
summary(catch_data)
```

## Fix Chinook Column

Use `mutate` to fix Chinook column

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))

summary(catch_clean)
```

Finding rows that were converted to NA

```{r}
i <- which(is.na(catch_clean$Chinook))
i
```

```{r}
catch_original[i, ]
```

Convert the "I" to 1

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook))

summary(catch_clean)
```


# Reshape Data

```{r}
catch_long <- catch_clean %>% 
  pivot_longer(
    cols = c(Chinook:Chum),
    names_to = "Species",
    values_to = "Count"
  )

head(catch_long)
```

```{r}
catch_wide <- catch_long %>% 
  pivot_wider(
    names_from = Year,
    values_from = Count
  )

head(catch_wide)
```

`rename` count column to `catch_thousands`

```{r}
catch_long <- catch_long %>% 
  mutate(catch = Count * 1000) %>% 
  select(-Count)
  
head(catch_long)  
```

# Summarize Data

```{r}
mean_region <- catch_long %>% 
  group_by(Region, Species) %>% 
  summarize(
    catch_mean = mean(catch),
    num_obs = n()
  )

mean_region
```

Calculate total catch per species
What is the most commonly caught species?

```{r}
total_species <- catch_long %>% 
  group_by(Species) %>% 
  summarize(catch_total = sum(catch)) %>% 
  arrange(desc(catch_total))

total_species
```

Using `filter`

```{r}
total_species_pre_1900 <- catch_long %>% 
  filter(Year < 1900) %>% 
  group_by(Species) %>% 
  summarize(catch_total = sum(catch)) %>% 
  arrange(desc(catch_total))

total_species_pre_1900
```


# Join to Region Table

Read in Region data table

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"), stringsAsFactors = FALSE) %>% 
  select(code, mgmtArea)

head(region_defs)
```

Join tables

```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)
```

# Separate and Unite

```{r}
site_codes <- data.frame(
  site = c(
    "HAW-100",
    "HAW-101",
    "HAW-102",
    "HAW-103",
    "HAW-100"
  ),
  stringsAsFactors = FALSE
)
```

Separate island and site number using `separate`

```{r}
site_codes_split <- site_codes %>% 
  separate(
    site, 
    into = c("island", "site_num"), 
    sep = "-",
    remove = FALSE
  )

site_codes_split
```

```{r}
site_codes_unite <- site_codes_split %>% 
  unite(col = site_code, island, site_num, sep = "_")

site_codes_unite
```

